---
title: "Introduction to the SICoxCURE package"
author: "Mohamed Yassir Errahmani, Mailis Amico, Ingrid Van Keilegom, Catherine Legrand"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the SICoxCURE package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

## Overview

The `SICoxCURE` package implements the **Single-Index/Cox (SIC) mixture
cure model** proposed by Amico, Van Keilegom and Legrand (2019).
The model writes the population survival function as:

$$S(t \mid \mathbf{x}, \mathbf{z}) =
  1 - p(\mathbf{x}) + p(\mathbf{x}) \, S_u(t \mid \mathbf{z})$$

where:

- $p(\mathbf{x}) = g(\boldsymbol{\gamma}^\top \mathbf{x})$ is the
  **incidence** (probability of being non-cured), estimated
  nonparametrically via Nadaraya-Watson kernel smoothing
- $S_u(t \mid \mathbf{z})$ is the **latency** (survival of non-cured
  subjects), modeled via a Cox proportional hazards model

## Installation

```{r install}
install.packages("SICoxCURE")
```

## Quick start

```{r simulate}
library(SICoxCURE)

# Simulate data from scenario 1 (logistic link)
set.seed(42)
n     <- 200
X     <- matrix(rnorm(n * 4), n, 4)
Z     <- matrix(rbinom(n, 1L, 0.6), n, 1)
p     <- 1 / (1 + exp(-X %*% c(1, -0.5, 0.3, -0.2)))
B     <- rbinom(n, 1L, p)
T0    <- ifelse(B == 1, rexp(n, 0.5), Inf)
C     <- rexp(n, 0.3)
Y     <- pmin(T0, C)
delta <- as.integer(T0 <= C & T0 < Inf)
dat   <- data.frame(Y=Y, delta=delta)
head(dat[, c("Y", "delta", "X1", "X2", "Z", "B", "p_true")])
```

## Fitting the SIC model

```{r fit}
X <- as.matrix(dat[, c("X1", "X2", "X3", "X4")])
Z <- as.matrix(dat[, "Z", drop = FALSE])

fit <- sic_fit(dat$Y, dat$delta, X, Z,
               h_step  = 0.05,
               verbose = 1)
summary(fit)
```

## Predictions

```{r predict}
# Predictions on training data
p_hat <- predict(fit, type = "incidence")

# Predictions on new data
pred <- predict(fit,
                newX   = X[1:10, ],
                newZ   = Z[1:10, ],
                type   = "all")
```

## Prediction error (train/test split)

```{r pred-error}
set.seed(1)
idx  <- sample(nrow(dat), 160)

fit2 <- sic_fit(dat$Y[idx], dat$delta[idx],
                X[idx, ], Z[idx, , drop = FALSE],
                h_step = 0.05, verbose = 0)

pe <- prediction_error(fit2,
                       Y_test     = dat$Y[-idx],
                       delta_test = dat$delta[-idx],
                       X_test     = X[-idx, ],
                       Z_test     = Z[-idx, , drop = FALSE])
cat("Prediction error:", round(pe, 3), "\n")
```

## Accuracy on simulated data

```{r ase}
ase <- compute_ase(fit$p_hat, dat$p_true)
cat("ASE:", round(ase, 5), "\n")
```

